---
title: "Build a model"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is an accompanying R Markdown document for the tidymodels get started article [Build a model](https://www.tidymodels.org/start/models/). 

If you ever get lost, you can click on the header to see the accompanying section in the online article.

## [Introduction](https://www.tidymodels.org/start/models/#intro)


```{r load-libs}
library(tidymodels)  # for the parsnip package, along with the rest of tidymodels

# Helper packages
library(readr)       # for importing data
```


## [The Sea Urchins Data](https://www.tidymodels.org/start/models/#data)


```{r data}
urchins <-
  # Data were assembled for a tutorial 
  # at https://www.flutterbys.com.au/stats/tut/tut7.5a.html
  read_csv("https://tidymodels.org/start/models/urchins.csv") %>% 
  # Change the names to be a little more verbose
  setNames(c("food_regime", "initial_volume", "width")) %>% 
  # Factors are very helpful for modeling, so we convert one column
  mutate(food_regime = factor(food_regime, levels = c("Initial", "Low", "High")))
```

Look at the data:

```{r}
urchins
```

Plot the data: 

```{r urchin-plot}
ggplot(urchins,
       aes(x = initial_volume, 
           y = width, 
           group = food_regime, 
           col = food_regime)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) +
  scale_color_viridis_d(option = "plasma", end = .7)
```

## [Build and fit a model](https://www.tidymodels.org/start/models/#build-model)

```{r lm-spec}
linear_reg() %>% 
  set_engine("lm")
?linear_reg()
```

Create model specification:

```{r}
lm_mod <- 
  linear_reg() %>% 
  set_engine("lm")
```

Fit model:

```{r lm-fit}
lm_fit <- 
  lm_mod %>% 
  fit(width ~ initial_volume * food_regime, data = urchins)
lm_fit
```

```{r lm-table}
tidy(lm_fit)
```

## [Use a model to predict](https://www.tidymodels.org/start/models/#predict-model)

New example data to predict:

```{r new-points}
new_points <- expand.grid(initial_volume = 20, 
                          food_regime = c("Initial", "Low", "High"))
new_points
```

Generate the mean body width values: 

```{r lm-pred-mean}
mean_pred <- predict(lm_fit, new_data = new_points)
mean_pred
```

Get confidence intervals and plot:

```{r lm-all-pred}
conf_int_pred <- predict(lm_fit, 
                         new_data = new_points, 
                         type = "conf_int")
conf_int_pred

# Now combine: 
plot_data <- 
  new_points %>% 
  bind_cols(mean_pred) %>% 
  bind_cols(conf_int_pred)

# and plot:
ggplot(plot_data, aes(x = food_regime)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, 
                    ymax = .pred_upper),
                width = .2) + 
  labs(y = "urchin size")
```

## [Model with a different engine](https://www.tidymodels.org/start/models/#new-engine)

Switch to Bayesian approach by simply changing your engine to **stan**: 

```{r go-stan, message = FALSE}
# set the prior distribution
prior_dist <- rstanarm::student_t(df = 1)

set.seed(123)

# make the parsnip model
bayes_mod <-   
  linear_reg() %>% 
  set_engine("stan", 
             prior_intercept = prior_dist, 
             prior = prior_dist) 

# train the model
bayes_fit <- 
  bayes_mod %>% 
  fit(width ~ initial_volume * food_regime, data = urchins)

print(bayes_fit, digits = 5)
```

To update the parameter table, the `tidy()` method is once again used: 

```{r tidy-stan}
tidy(bayes_fit, intervals = TRUE)
```

Get your predictions easily without changing the syntax:

```{r stan-pred}
bayes_plot_data <- 
  new_points %>% 
  bind_cols(predict(bayes_fit, new_data = new_points)) %>% 
  bind_cols(predict(bayes_fit, new_data = new_points, type = "conf_int"))

ggplot(bayes_plot_data, aes(x = food_regime)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, ymax = .pred_upper), width = .2) + 
  labs(y = "urchin size") + 
  ggtitle("Bayesian model with t(1) prior distribution")
```


## [Why does it work that way?](https://www.tidymodels.org/start/models/#why)

Use the pipe to pass around the _data_ in the **tidyverse**:

```{r tidy-data}
urchins %>% 
  group_by(food_regime) %>% 
  summarize(med_vol = median(initial_volume))
```

Use the pipe to pass around the _model object_ with **tidymodels**

```{r tidy-model, eval = FALSE}
bayes_mod %>% 
  fit(width ~ initial_volume * food_regime, data = urchins)
```

This may seem jarring if you have used dplyr a lot, but it is extremely similar to how ggplot2 operates:

```{r eval=FALSE}
ggplot(urchins,
       aes(initial_volume, width)) +      # returns a ggplot object 
  geom_jitter() +                         # same
  geom_smooth(method = lm, se = FALSE) +  # same                    
  labs(x = "Volume", y = "Width")         # etc
```

